Infinite Cellular Automata Running Virtual 6502 Processors

### 2018-07-02 paper.tex
CA

The cellular automaton (CA) is a $L \times L \times L$ array of cells.
Cells are indexed $(x,y,z$) where
$\alpha \leq x \leq \alpha+L$,
$\beta \leq y \leq \beta+L$,
$\gamma \leq z \leq \gamma+L$
and $(\alpha,\beta,\gamma)$ are integers representing global coordinates of this CA.

@@

Cells in the CA are updated in a pseudorandom order.
This can be any order that visits all cells uniformly;
a Mersenne Twister is suitable for this,
and also for the generation of the random variable {\tt RUNTIME} used to constrain the update (see below).

@@

On termination, the transient address ranges in the memory map
(including the stack)
are discarded.
The neighborhood memory map (\hexrange{4000}{7FFF}, Table~\ref{tab:MooreNeighborhood})
is mapped back to the cellular automata storage.

@@

\begin{table}
\mooremap
\caption{
  \label{tab:MooreNeighborhood}
  Memory map for the 6502 virtual machine of cell $(x,y,z)$,
  using the 26-cell three-dimensional Moore neighborhood.
  All address ranges not defined in this table are read-only and zero; writing to them has no effect.
}
\end{table}

\begin{table}
\vonneumannmap
\caption{
  \label{tab:VonNeumannNeighborhood}
  Memory map for the 6502 virtual machine of cell $(x,y,z)$,
  using the 6-cell three-dimensional von Neumann neighborhood.
}
\end{table}

>>>>

The cellular automaton (CA) is a $L \times L \times L$ array of cells.
Cells are indexed $(x,y,z$) where
$\alpha_x \leq x < \alpha_x+L$,
$\alpha_y \leq y < \alpha_y+L$,
$\alpha_z \leq z < \alpha_z+L$
and $(\alpha_x,\alpha_y,\alpha_z)$ are integers representing global coordinates of this CA.

@@

An {\bf update} of the CA consists of updating each cell in the CA in a pseudorandom order.
This can be any order that visits all cells uniformly;
a Mersenne Twister is suitable for this,
and also for the generation of the random variable {\tt RUNTIME} used to constrain the update (see below).

@@

On termination, the transient address ranges in the memory map
(including the stack)
are discarded.
The neighborhood memory map (\hexrange{4000}{7FFF}, Table~\ref{tab:MooreNeighborhood})
is copied back to the cellular automata storage.

\section{Adjacencies}

Consider those cell coordinates $(x,y,z)$ that are outside the CA,
e.g. $(\alpha_x-1,\alpha_y,\alpha_z)$ or  $(\alpha_x,\alpha_y,\alpha_z + L)$.
Some of these, specifically those in the Moore neighborhood of cells that are in the CA,
are a relevant initial condition for the machine,
since they map to bytes in its addressable memory.

@@

\begin{table}
\mooremap
\caption{
  \label{tab:MooreNeighborhood}
  Memory map for the 6502 virtual machine of cell $(x,y,z)$,
  using the 26-cell three-dimensional Moore neighborhood.
  All address ranges not defined in this table are read-only and zero; writing to them has no effect.
  Pages mapping to cells outside the CA (e.g. $x<\alpha_x$ or $x \geq \alpha_x+L$, or similarly for $y,z$)
  are zero when read (unless otherwise specified),
  and writing to them has no effect.
}
\end{table}

\begin{table}
\vonneumannmap
\caption{
  \label{tab:VonNeumannNeighborhood}
  Memory map for the 6502 virtual machine of cell $(x,y,z)$,
  using the 6-cell three-dimensional von Neumann neighborhood.
  This memory map may be considered a subset of the memory map in Table~\ref{tab:MooreNeighborhood}.
}
\end{table}
